\documentclass[10pt,a4paper]{report}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\author{Sam Stern (s1134468)}
\title{Machine Learning with Domestic Energy Use Data}
\begin{document}
\maketitle

\begin{abstract}
As part of the UK Government's insentive to reduce the Nation's energy consumption, smart meters are being rolled out to households and small businesses accross the UK. In this project aims to assess some of the security risks associated with gathering data relating to a households energy consumption.
\end{abstract}

\tableofcontents

\input{tex/introduction}


\chapter{Data}


\chapter{Feature Exploration and Extraction}

According to \cite{Beckel 2}, possible features that are interesting for classification of househods based on energy consumption are: consumption figures, ratios, temporal properties, and statistical properties. Consumption figures are the average, maximum and minimum energy consumption over some time period. Ratios are features that calculate the ratio between consumption figures and can capture relevant patterns that occure through different time intervals. Temporal features capture the first (or last) time some event takes place which or at what time the daily maximum occures. Finally, statistical properties such as variance, give insight into the consumption curve (for example how a households energy consumption correlates with itself.
\section{Building Features from Energy Readings}





\begin{thebibliography}{9}

\bibitem{Beckel 1} Christian Beckel, Leyna Sadamori, and Silvia Santini. Automatic socio-economic classification of households using electricity consumption data. In {\em Proceedings of the Fourth International Conference on Future Energy Systems}, e-Energy `13, pages 75–86, New York, NY, USA, 2013. ACM.

\bibitem{Beckel 2} Christian Beckel, Leyna Sadamori, and Silvia Santini. Towards automatic classifi- cation of private households using electricity consumption data. In {\em Proceedings of the Fourth ACM Workshop on Embedded Sensing Systems for Energy-Efficiency in Buildings}, BuildSys `12, pages 169–176, New York, NY, USA, 2012. ACM.

\bibitem{OfGEM} Office of Gas and Electricity Markets (OfGEM). {\em Transition to smart meters}. {\tt https://www.ofgem.gov.uk/electricity/retail-market/metering/ transition-smart-meters}.

\bibitem{Vasc} Jorge Vasconcelos. Survey of regulatory and technological developments concern- ing smart metering in the european union electricity market. {\tt http://hdl.handle. net/1814/9267, 2008.}
\end{thebibliography}
\end{document}
